A Method for Regularization of Evolutionary Polynomial Regressions

### 2013-12-22 texts/submission_02-asc/paper_spfga-review.tex
Pre-submit version.

The performance of each method is evaluated on several common datasets. From each dataset 70\% of the observations are reserved for training purposes and the remaining observations used to estimate the \ac{rmse}. To enhance the robustness of results this process is repeated $25$~times, each time with different shuffling of the samples in the train and test sets. Some datasets with attribute values of different magnitudes have a pre-processing scaling transformation. The box plots in figures \ref{artificial_dataset1_lambda1.0} and \ref{fig:four.datasets.summary} resume the test set error distributions over these different runs.

@@

% Of all the regression methods considered, the proposed method is the one that shows, with one exception, the lowest error. Only Random Forest outperforms \ac{EPR} systematically (it also outperforms all the other regression algorithms). The single non-ensemble exception --- besides the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{EPR} has comparable results. This is evidence that applying standard genetic operators for polynomial model searching is a viable tool for regression purposes.

Of the regression methods considered, \ac{SVM} achieved the best results in 3 out of 4 datasets. However, \ac{SVM} was tuned for each particular dataset while \ac{EPRR} was executed with its parameters preset. Even so, \ac{EPRR} had the best error in the \textsc{Abalone} dataset, and achieved competitive results in the \textsc{Housing} and \textsc{Auto MPG} datasets.

Comparing \ac{EPR} and \ac{EPRR} -- which is the main article's topic -- the regularized version achieved much better results at \textsc{Abalone} and especially \textsc{Kinematics}. The \textsc{Housing} dataset improved errors when compared with \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance. Only in the \textsc{Auto MPG} dataset \ac{EPR} achieved better results, even if not that different from \ac{EPRR}.

For complexity considerations \ac{EPR} and \ac{EPRR} demands some processing time. On a quad-core computer, processing the \textsc{Kinematics} dataset (with 8k~observations) takes approximately 5~minutes. The processing time can probably be speeded between one to two orders in magnitude if the process is implemented in a low level programming language like \texttt{C++}. However, speed optimization was not the focus of this article.

>>>>

The performance of each method is evaluated on several common datasets. From each dataset 70\% of the observations are reserved for training purposes and the remaining observations used to estimate the \ac{rmse}. To enhance the robustness of results this process is repeated $25$~times, each time with a different shuffling of the samples in the train and test sets. Some datasets with attribute values of different magnitudes have a pre-processing scaling transformation. The box plots in figures \ref{artificial_dataset1_lambda1.0} and \ref{fig:four.datasets.summary} resume the test set error distributions over these different runs.

@@

Of the regression methods considered \ac{SVM} achieves the best results in 3 out of 4 datasets. However \ac{SVM} and Conditional Inference Trees are pre-trained, with parameters tuned for each particular dataset unlike \ac{EPRR}. Even so it is the best estimator for the \textsc{Abalone} dataset and in the remaining datasets it outperforms most of the other estimators.

Comparing \ac{EPR} and \ac{EPRR} --- the main article's topic --- the regularized version achieved much better results at \textsc{Abalone} and especially \textsc{Kinematics}. The \textsc{Housing} dataset improved errors when compared with \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance. Only in the \textsc{Auto MPG} dataset \ac{EPR} achieved better results, even if not that different from \ac{EPRR}.

For complexity considerations \ac{EPR} and \ac{EPRR} demand some processing time. On a quad-core computer, processing the \textsc{Kinematics} dataset (with 8k~observations) takes approximately 5~minutes. The processing time can probably be speeded between one to two orders in magnitude if the process is implemented with computational speed in mind. However, errospeed optimization is not the focus of this article.
